---
title: "Project2"
author: "Yiting Zhang"
date: "2/13/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}



# Clear all variables and prior sessions
rm(list=ls(all=TRUE))

# Load libraries
library(fma)
library(tseries)
library(forecast)
library(fpp3)
library(tseries)
library(seasonal)
library(tsibble)
library(dyn)
library(ggplot2)
library(Hmisc)
library(latex2exp)
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library(xtable)
library(stats)
library(dyn)
library(plotrix)
library(strucchange)
library(quantmod)
```

```{r}

google =read.csv("GOOG_data.csv")

fb =read.csv("FB_data.csv")

return =read.csv("Return.csv")

google <- google[-2450,]

```

```{r}

ts_google <- ts(google$Adj.Close, start=c(2012,5),frequency=365)

ts_fb <- ts(fb$Adj.Close, start=c(2012,5),frequency=365)

```


## j
Compare your forecast from (i) to the 12-steps ahead forecasts from ARIMA, Holt-Winters,
and ETS models. Which model performs best in terms of MAPE?


## k
Combine the four forecasts and comment on the MAPE from this forecasts vs., the individual
ones.

## l
Fit an appropriate VAR model using your two variables. Make sure to show the relevant plots
and discuss your results from the fit.

According to the plot, returns of goodle and facebook all show smean reverting nature. 

```{r}
plot(ts_google)
plot(ts_fb)
```

```{r}
library(vars)

VARselect(data.frame(ts_google, ts_fb))

```

The results from the VAR model shows the most significance for the lower lags and lags from 5-8. We need impulse response to have further analysis.
```{r}

model <- VAR(data.frame(ts_google, ts_fb), 9)
summary(model)

```


## m
Compute, plot, and interpret the respective impulse response functions.

The impulse response from google shows a correlation of facebook with respect to the fluctuation of google, while opposite does not show a significant response of google to facebook. 

```{r}

plot(irf(model), plot.type = "multiple", names = "ts_google")

plot(irf(model), plot.type = "multiple", names = "ts_fb")


```

## n
Perform a Granger-Causality test on your variables and discuss your results from the test.

The Granger test shows that facebook stock price is dependent on google while the converse is not true.

```{r}
# Picks order 9
grangertest(ts_fb ~ ts_google, order = 9)

grangertest(ts_google ~ ts_fb, order = 9)

```

## o
Use your VAR model to forecast 12-steps ahead. Your forecast should include the respective
error bands. Comment on the differences between the VAR forecast and the other ones
obtained using the different methods.

```{r}

prd <- predict(model, n.ahead = 12, ci = 0.95, dumvar = NULL)
print(prd)
plot(prd)
plot(prd, xlim=c(2400,2460))

```